A Technique for Inter-Retailer Recommendations

### 2014-06-12 paper2/paper.tex
Small edits

\subsubsection*{Concept Space Mapping}
Using random walks to simulate user data, followed by the application of LSI, is
indeed an effective way to encode information from a recommendation graph into a
latent feature space. But what is the relationship between latent feature
vectors derived from disconnected graph partitions? In principle, there is no
guarantee of any relationship between such concept spaces in the same way that
there is no guarantee of any relationship between two arbitrary, disconnected
recommendation graphs. Nonetheless, our hypothesis is that a relationship
between these concept spaces often will exist due to the effectiveness of the
underlying CF techniques that generate the graphs to capture meaningful
item-item associations. Put another way, we believe that the most fundamental
latent dimensions along which products in a category such as Women's swimsuits
vary will be common to many swimsuit catalogues. If this is indeed true, then it
might be possible to learn a mapping between these concept spaces.

If a mapping between the concept spaces associated with two catalogues were
known, we could translate all items into a common space and then choose
inter-catalogue edges between item pairs that are nearest to one another in
concept space. To learn these mapping, we must revisit the notion of content
similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
an item $i$'s tf-idf vector $\vec{\tau}(i)$ by the elements in its latent
feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
super-tf-idf} vector for each concept for each catalogue. Formally, we say that
a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
\begin{align}
\vec{s}(j,C) = \sum_{i \in C}{v(i)_j \cdot \vec{\tau}(i)},
\end{align}
where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
feature vector. Note that since the components $v(i)_j$ may be positive or
negative so too can the elements of a concept's super-tf-idf vector, in
constrast to the elements of an individual item's tf-idf vector which are
necessarily greater than or equal to zero. Also note that we compute both
$\vec{s}_1$ and $\vec{s}_2$, which correpond to unigram and bigram super-tf-idf
vectors, respectively, for each concept. These super-tf-idf vectors then act as
proxies for the concept.

@@

It is important to note that the transformation matrix $A$ is not orthogonal, and
therefore, does not preserve vector norms. We address this issue by normalizing
all feature vectors, both foreign and domestic, in the concept space following
the transformation. This radially projects all items onto the surface of sphere
within the 16-dimensional concept space. Now when ranking an item's neighbors by
distance, we are actually ranking neighbors by angular separation. That is, we
are ranking neighbors by their cosine similarity in concept space. After
experimenting with various normalization schemes we find that this approach
works best.

>>>>

\subsubsection*{Concept Space Mapping}
Using random walks to simulate user data, followed by the application of LSI, is
indeed an effective way to encode information from a recommendation graph into a
latent feature space. But what is the relationship between latent feature
vectors derived from disconnected graph partitions? In principle, there is no
guarantee of any relationship between such concept spaces in the same way that
there is no guarantee of any relationship between two arbitrary, disconnected
recommendation graphs. Nonetheless, our hypothesis is that a relationship
between these concept spaces will often exist due to the effectiveness of the
underlying CF techniques that generate the graphs to capture meaningful
item-item associations. Put another way, we believe that the most fundamental
latent dimensions along which products in a category such as Women's swimsuits
vary will be common to many swimsuit catalogues. If this is indeed true, then it
might be possible to learn a mapping between these concept spaces.

If a mapping between the concept spaces associated with two catalogues were
known, we could translate all items into a common space and then choose
inter-catalogue edges between item pairs that are nearest to one another in
concept space. To learn these mappings, we must revisit the notion of content
similarity. Once again, we utilize tf-idf vectors. This time, however, we weight
an item $i$'s tf-idf vector $\vec{\tau}(i)$ by the elements in its latent
feature vector $\vec{v}(i)$. In this way, we construct a dense {\em
super-tf-idf} vector for each concept of each catalogue. Formally, we say that
a concept $j$ of a catalogue $C$ is represented by the super-tf-idf vector
\begin{align}
\vec{s}(j,C) = \sum_{i \in C}{v(i)_j \cdot \vec{\tau}(i)},
\end{align}
where $v(i)_j$ is the $j$\textsuperscript{th} component of item $i$'s latent
feature vector. Note that since the components $v(i)_j$ may be positive or
negative so too can the elements of a concept's super-tf-idf vector, in
constrast to the elements of an individual item's tf-idf vector which are
necessarily greater than or equal to zero. Also note that we compute both
$\vec{s}_1$ and $\vec{s}_2$, which correpond to unigram and bigram super-tf-idf
vectors, respectively, for each concept. These super-tf-idf vectors then act as
proxies for the concept.

@@

It is important to note that the transformation matrix $A$ is not orthogonal, and
therefore, does not preserve vector norms. We address this issue by normalizing
all feature vectors, both foreign and domestic, in the concept space followingj
the transformation. This radially projects all items onto the surface of a
hypersphere in the 16-dimensional concept space. Now, when ranking an item's
neighbors by distance, we are actually ranking neighbors by angular separation.
That is, we are ranking neighbors by their cosine similarity in concept space.
After experimenting with various normalization schemes we find that this
approach works best.
